---
title: "Homework6"
author: "Anna Schmoker"
date: "2/21/2018"
output: 
  html_document: 
    highlight: tango
    theme: cerulean
---
## Reading actual data

```{r}
z <- read.table("TestDataSet_CrkLSH2screen_Score.csv",header=TRUE,sep=",", stringsAsFactors=FALSE)
str(z)
summary(z)
head(z)
```

## Opening Libraries

```{r}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```

## Plotting my data

```{r}
p1 <- ggplot(data=z, aes(x=Score, y=..density..)) +
  geom_histogram(color="turquoise4",fill="turquoise",size=0.2) 
print(p1)
```

## Add an empirical density curve

```{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

## Get maximum likelihood parameters

```{r}
normPars <- fitdistr(z$Score,"normal")
print(normPars)
```

```{r}
str(normPars)
```

```{r}
normPars$estimate["mean"] # note structure of getting a named attribute
```

## Plot normal probability density

```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$Score),len=length(z$Score))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$Score), args = list(mean = meanML, sd = sdML))
 p1 + stat
```

Not a great fit...

## Plot exponential probability density

```{r}
expoPars <- fitdistr(z$Score,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$Score), args = list(rate=rateML))
 p1 + stat + stat2
```

...better, in a way

## Plot uniform probability density

```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$Score), args = list(min=min(z$Score), max=max(z$Score)))
 p1 + stat + stat2 + stat3
```


## Plot gamma probability density

```{r}
z2 <- z[z$Score>0,]
gammaPars <- fitdistr(z2$Score,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z2$Score), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```


## Plot beta probability density

```{r}
pSpecial <- ggplot(data=z, aes(x=Score/(max(Score + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z2$Score/max(z2$Score + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z2$Score), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

```{r}
normPars$loglik
gammaPars$loglik
expoPars$loglik
```

Although not a great fit, the exponential fit wins today

## Extracting the maximum likelihood params from the exponential fit

```{r}
expoPars
```

## Simulating new data -- Exponential

```{r}
x <- rexp(n=8887,rate=2.52)
x <- data.frame(1:8887,x)
names(x) <- list("density","Score")
str(x)
```

```{r}
summary(x$Score)
```

```{r}
p2 <- ggplot(data=x, aes(x=Score, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk") 
p2 + stat2
p1 + stat2
```

# How do the plots compare?

Not very well. My data was somewhat bimodal, and not a great fit for any of these distributions. I think this is due to the nature of the data, itself, which I don't think is necessarily able to be simulated. I think about other data types that I have collected that may be better to use for this exercise and return to it.
